Equalizer\\Programming and User Guide

### 2010-06-03 Developer/ProgrammingGuide/paper.tex
PG update
[ ] May break build
[ ] Breaks existing applications
[ ] Bugfix
[ ] New Feature
[ ] Cleanup
[ ] Optimization
[x] Documentation

\date{
  \vspace{14cm}
\begin{textblock}{}(0,4.5)
  \hspace{-1cm}\includegraphics[width=22cm]{images/teaser.pdf}
\end{textblock} {\Large The official reference for developing and
  deploying parallel, scalable OpenGL\texttrademark\ applications based on
  the Equalizer parallel rendering framework}\\\vspace{1cm}
Version 1.7.4 for Equalizer 0.9.1+svn\\[\medskipamount]
%  \today
}

@@

\clearpage
\pagenumbering{roman}
\pagestyle{headings}
\tableofcontents
\clearpage
\listoffigures \vfill
\section*{Revision History}
{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Revision for Equalizer 0.5\\
    1.4     & Nov 25, 2008 & Revision for Equalizer 0.6\\
    1.6     & Aug 07, 2009 & Revision for Equalizer 0.9\\
    1.7.0   & Sep 09, 2009 & Updates based on hardcopy print\\
    1.7.1   & Oct 20, 2009 & Updates for new eq::GLWindow\\
    1.7.2   & Nov 18, 2009 & Add \sref{sSubpixel} and \sref{sFSAA} for Subpixel compounds\\
    1.7.3   & Feb 02, 2010 & Revision for Equalizer 0.9.1\\
    1.7.4   & Feb 22, 2010 & Add \sref{sSlaveCommit}\\
    \hline
  \end{tabularx}}
\clearpage

@@

>>>>

\date{
  \vspace{14cm}
\begin{textblock}{}(0,4.5)
  \hspace{-1cm}\includegraphics[width=22cm]{images/teaser.pdf}
\end{textblock} {\Large The official reference for developing and
  deploying parallel, scalable OpenGL\texttrademark\ applications based on
  the Equalizer parallel rendering framework}\\\vspace{1cm}
Version 1.7.5 for Equalizer 0.9.1+svn\\[\medskipamount]
%  \today
}

@@

\clearpage
\pagenumbering{roman}
\pagestyle{headings}
\tableofcontents
%\clearpage
\listoffigures \vfill
\section*{Revision History}
{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Revision for Equalizer 0.5\\
    1.4     & Nov 25, 2008 & Revision for Equalizer 0.6\\
    1.6     & Aug 07, 2009 & Revision for Equalizer 0.9\\
    1.7.0   & Sep 09, 2009 & Updates based on hardcopy print\\
    1.7.1   & Oct 20, 2009 & Updates for new eq::GLWindow\\
    1.7.2   & Nov 18, 2009 & Add \sref{sSubpixel} and \sref{sFSAA} for Subpixel compounds\\
    1.7.3   & Feb 02, 2010 & Revision for Equalizer 0.9.1\\
    1.7.4   & Feb 22, 2010 & Add \sref{sSlaveCommit}\\
    1.7.5   & Jun 03, 2010 & Add \sref{sClusterSetup}\\
    \hline
  \end{tabularx}}
\clearpage

@@

\section{\label{sClusterSetup}Setting up a Visualization Cluster}
This section covers the setup of a visualization cluster to run Equalizer
applications. It does not cover basic cluster management and driver
installation. A prerequisite to the following steps is a pre-installed cluster,
including network and graphics card drivers.

\subsection{Name Resolution}
Pay attention that your name resolution works properly, that is, the output of
the hostname command should resolve to the same, non-local IP address on all
machines. If this is not the case, you will have to provide the public IP
address for all processes, in the following way:
\begin{description}
\item[Server] Specify the server IP as the hostname field in the connection
  description in the server section.
\item[Application] Start the application using the IP as the listening port,
  e.g., \textsf{./build/Linux/bin/eqPly -- --eq-listen 192.168.0.1}
\item[Render Clients] Specify the client IP's as the hostname field in the
  connection description of each node section.
\end{description}

\subsection{The Server}
The server may be started as a separate process or within the application
process. If it is started separately, simply provide the desired configuration
file as a parameter. It will listen on all network addresses, unless other
connection parameters are specified in the configuration file. If the
server is started within the application process, using the \textsf{--eq-config}
parameter, you will have to specify a connection description for the server in
the configuration file. Application-local servers do not create a listening
socket by default for security reasons.

\subsection{Starting Render Clients}
Cluster configurations use multiple Equalizer nodes, where each node represents
a separate process, typically on a different machine. These node processes have
to be started for a visualization session, and need to communicate with each
other.

Equalizer supports pre-launched render clients and automatic render client
launching, if configured properly. Pre-launched render clients are started
manually or by an external script on a pre-defined address. Auto-launched render
client are started and stopped by the Equalizer server on demand.

The two mechanism can co-exist. The server will first try to connect a
pre-launched render client on the given connection descriptions for each
node. If this fails, he will try to auto-launch the render client. If the render
client does not connect back to the server within a certain timeout (default one
minute), a failure to initialize the configuration is reported back to the
application.

\subsubsection{Pre-launched Render Clients}
Pre-launched render clients are useful if setting up auto-launching is too
time-consuming, e.g., on Windows which does not have a standard remote login
procedure. The following steps are to be taken to use pre-launched render
clients:

\begin{itemize}
\item Set the \textbf{connection} hostname and port parameters of each node in
  the configuration file.
\item Start the \textbf{render clients} using the parameters
  \textsf{--eq-client} and \textsf{--eq-listen}, e.g.,
  \textsf{./build/Linux/bin/eqPly -- --eq-client --eq-listen
  192.168.0.2:1234}. Pay attention to use the same connection parameters as in
    the configuration file.
\item Start the \textbf{application}. If the server is running on the same
  machine and user as the application, the application will connect to it
  automatically. Otherwise use the \textsf{--eq-server} parameter to specify the
  server address.
\end{itemize}

The render clients will automatically exit when the config is closed. The
\textsf{eqPly} example application implements the option \textsf{-r} to keep the
render client processes resident.

\subsubsection{Auto-launched Render Clients}
To automatically launch the render clients, the server needs to know the name of
the render client and the command to launch them without user interaction.

The name of the render client is automatically set to the name of the
application executable. This may be changed programmatically by the
application. Normally it suffices to install the application in the same
directory on all machines, ideally using a shared file system.

The default launch command is set to ssh, which is the most common solution for
remote logins. To allow the server to launch the render clients without user
interaction, password-less ssh needs to be set up. Please refer to the ssh
documentation (cf. \textsf{ssh-keygen} and \textsf{~/.ssh/authorised\_keys}) and
verify the functionality by logging in to each machine from the host running the
server.

\subsection{Debugging}
If your configuration does not work, simplify it as much as possible
first. Normally this means that there is one server, application and render
client. Failure to launch a cluster configuration often is caused by one of the
following reasons:
\begin{itemize}
\item A firewall is blocking network connections.
\item The render client can't access the GPU's on the remote host. Set up your
  X-Server or application rights correctly. Log into the remote machine using
  the launch command and try to execute a simple application, e.g.,
  \textsf{glxinfo -display :0.1}.
\item The server does not find the pre-launched render client. Verify that the
  client is listening on the correct IP and port, and that this IP and port are
  reachable from the server host.
\item The server can not launch a render client. Check the server log for the
  launch command used, and try to execute a simple application from the server
  host using this launch command. It should run without user interaction. Check
  that the render client is installed in the correct path. Pay attention to the
  launch command quotes used to separate arguments on Windows. Check that the
  same software version, including Equalizer, are installed on all machines.
\item A client can't connect back to the application. Check the client log, this
  is typically caused by a misconfigured host name resolution.
\end{itemize}
